Constraint Grammar as a SAT problem

### 2015-06-01 paper/nodalida2015.tex
small fixes

\subsection{Rules can be unordered}

@@

When conducting our tests, we have been able to construct small
grammars (\textless20 rules) which, when run with the unordered scheme,
disambiguate more accurately than the same grammar run with VISL
CG-3. However, when testing with larger grammars that are written with the
traditional CG in mind, SAT-CG loses with both ordering strategies, but emulating
order fares better.
In the following sections, we report both execution time and
performance against state of the art.

@@

We took a manually tagged
corpus\footnote{https://svn.code.sf.net/p/apertium/svn/branches/apertium-swpost/apertium-en-es/es-tagger-data/es.tagged}
containing 21865 words of Spanish news text, 
and a small constraint grammar\footnote{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}, produced independently of the authors.
% We ignored substitute rules, but kept subreadings and unification, even though our implementation doesn't handle them.
We kept only select and remove rules, after which the grammar had 261 rules.
With this setup, we took the text of the tagged corpus with all
ambiguities, and ran both SAT-CG and VISL CG-3 with the same grammar. 
Treating the corpus as the gold standard, the disambiguation by  VISL
CG-3 achieves F-score of 82.6 \%, ordered
SAT-CG 81.5 \%  and unordered SAT-CG 79.2 \%. 
For comparison, with the best handcrafted grammar, unordered SAT-CG
achieves 85.3 \% and VISL CG-3 85.9 \%.
Similar patterns were observed with small (\textless{}20) rule sets
written by the authors; depending on the subset, SAT-CG and VISL CG-3 had
a difference of at most $\Mypm$ 1.5 \%. 
Introducing rules one by one up to 19, the
performance improved in a very similar rate, with less than 0.5 \%
difference between the systems at each new rule.
We did not evaluate on other languages or text genres, due to lack of suitable test data.

@@

We tested the performance by parsing Don Quijote (384,155 words) with
the same Spanish grammar. Table~\ref{table:time} shows execution times compared to VISL
CG-3 for the same text and rules. 
The number of rules in the grammar is more significant than the word
count; this holds also for the performance of VISL CG-3.
From the SAT solving side, maximisation is the most costly operation; 
emulating order (SAT-CG\textsubscript{o}) is slightly faster, because maximisation problems are smaller.
SAT is not Parsing Don Quijote (384,155 words) with the rule set of 261 rules,
the function was called 147,253 times, and with 19 rules, 132,255 times.
%However, with both rule sets, half of the sentences in Don Quijote
%needed less than 6 calls of maximise, and 75 \% of the sentences
%needed less than 14.
The difference in the execution times suggests that there are other
reasons for the worse performance---this is to be expected,
SAT-CG is currently just a naive proof-of-concept implementation with no optimisations.

@@

\section{Applications and Future Work}
\label{sec:apps}
As described in Section~\ref{sec:eval}, our implementation is not
competetive with the state of the art. It is more fruitful to focus on
things where the SAT-based representation can make an impact.
One such thing is grammar analysis.
There has been work on automatic tuning of hand-written CGs \todo{cite
Eckhard}, but no tools which would search for inconsistencies or
unoptimal design in order to point it out to the grammar writer.

Traditional CG rules are applied deterministically and
sequentially; at the execution of rule, there is no way to
go back to earlier rules and undo them. Thus there is no way to know
what part of the input comes from the morphological analyser and what
part from applying previous rules.

A conflict can be defined as 
a set of two or more rules, such that applying the first makes the
next rules impossible to apply, regardless of input.
As a simple example, these rules could be

@@

Maximisation-based conflict solving (see Section 3) is not feasible for this task. 
The whole definition of conflicts depends on ordering, and it would
not differentiate between the case where \texttt{SELECT verb} comes
before \texttt{REMOVE verb IF (-1 det)} or after.
On the other hand, a formalism such as FSIG would benefit from the
maximisation-based technique in conflict handling.

>>>>

\subsection{Rules can be unordered}
\label{sec:unord}

@@

In the following sections, we report both execution time and
performance against state of the art.

@@

We took a manually tagged
corpus\footnote{https://svn.code.sf.net/p/apertium/svn/branches/apertium-swpost/apertium-en-es/es-tagger-data/es.tagged}
containing approximately 22,000 words of Spanish news text, 
and a small constraint grammar\footnote{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}, produced independently of the authors.
% We ignored substitute rules, but kept subreadings and unification, even though our implementation doesn't handle them.
We kept only select and remove rules, after which the grammar had 261 rules.
With this setup, we produced an ambiguous version of the tagged
corpus, and ran both SAT-CG and VISL CG-3 on it.
Treating the original corpus as the gold standard, the disambiguation
by  VISL CG-3 achieves F-score of 82.6 \%, ordered SAT-CG 81.5 \%  and
unordered SAT-CG 79.2 \%. 
We did not test with other languages or text genres due to lack of
available gold standard.
%This result suggests that that when running grammars that are written
%with the traditional CG in mind, SAT-CG loses with both ordering
%strategies, but emulating order fares better.

We also wanted to test the hypothesis that SAT-CG needs different
kinds of rules to perform well. We took the base grammar and started
handcrafting rules; with our best performing and most concise
grammar\footnote{https://github.com/inariksit/cgsat/blob/master/data/spa\_smallset.rlx}
of just 19 rules, unordered SAT-CG achieves 85.3 \% and VISL CG-3 85.9
\%. This experiment is naturally very small and might be explained by
overfitting or mere chance, but it seems to indicate that rules that
work well with SAT-CG, are also good for traditional CG.
% It was possible to attain a rule set for which the unordered SAT-CG
% works better (\textless0.5 \%), but it required trying
% subsequences of the 19-rule set in a brute force manner---this hardly
% presents any real life use case.

% Similar patterns were observed with small (\textless{}20) rule sets
% written by the authors; depending on the subset, SAT-CG and VISL CG-3 had
% a difference of at most $\Mypm$ 1.5 \%. 
% Introducing rules one by one up to 19, the
% performance improved in a very similar rate, with less than 0.5 \%
% difference between the systems at each new rule.
% We did not evaluate on other languages or text genres, due to lack of suitable test data.

@@

We tested the performance by parsing Don Quijote (384,155 words) with
the same Spanish grammars as in previous experiment. Table~\ref{table:time} shows execution times compared to VISL
CG-3 for the same text and rules. 
The number of rules in the grammar is more significant than the word
count; this holds also for the performance of VISL CG-3.
From the SAT solving side, maximisation is the most costly operation; 
emulating order (SAT-CG\textsubscript{o}) is slightly faster, because maximisation problems are smaller.
In any case, SAT does not seem to be the bottleneck: with 261 rules,
the maximisation function was called 147,253 times, and with 19 rules,
132,255 times. 
%However, with both rule sets, half of the sentences in Don Quijote
%needed less than 6 calls of maximise, and 75 \% of the sentences
%needed less than 14.
The difference in the execution times suggests that there are other
reasons for the worse performance---this is to be expected,
SAT-CG is currently just a naive proof-of-concept implementation with no optimisations.

@@

\section{Applications and Future Work}
\label{sec:apps}
As described in Section~\ref{sec:eval}, our implementation is not
competetive with the state of the art. It is more fruitful to focus on
things where the SAT-based representation can make an impact.
One such thing is grammar analysis\footnote{We thank Eckhard Bick for
  the suggestion.}.
There has been work on automatic tuning of hand-written CGs
\cite{bick2013tuning}, but to our knowledge, no tools which would
search for inconsistencies or unoptimal design in order to point it out to the grammar writer.

The sequential application of traditional CG rules is good for
performance and transparency. As soon as a rule applies, the analyses
are removed from the input, and next rules get the modified sentence
as their input. At the execution of rule, there is no way to
go back to earlier rules and undo them. Thus there is no way to know
what part of the input comes from the morphological analyser and what
part from applying previous rules.

A conflict can be defined as 
a set of two or more rules, such that applying the first makes the
next rules impossible to apply, regardless of input.
We can reuse the example from Section~\ref{sec:unord}:

@@

Maximisation-based conflict solving (see Section~\ref{sec:unord}) is not feasible for this task. 
The whole definition of conflicts depends on ordering, and it would
not differentiate between the case where \texttt{SELECT verb} comes
before \texttt{REMOVE verb IF (-1 det)} or after.
On the other hand, a formalism such as FSIG would benefit from the
maximisation-based technique in conflict handling.
